```{r}

library(lattice)
require(grDevices)

```

# Title

## Synopsis (no more than 10 sentences)

## Data Processing (at least one figure but no more than 3)

```{r unzipAndLoad, cache=TRUE}

fullData <- read.csv("StormData.csv.bz2", strip.white=TRUE)

```

```{r restrictColumns, cache=TRUE} 

columnsToInclude <- c("EVTYPE", "FATALITIES", "INJURIES", "PROPDMG", "PROPDMGEXP", "CROPDMG", "CROPDMGEXP", "WFO", "REFNUM")

simpleData <- fullData[, columnsToInclude]

```

```{r fatalitiesPlusInjuries}

simpleData[,"CASUALTIES"] <- simpleData$INJURIES + simpleData$INJURIES

casualtiesVsType <- aggregate(simpleData$CASUALTIES, by=list(simpleData$EVTYPE), FUN=mean, na.rm=TRUE)

names(casualtiesVsType) = c("EVTYPE", "CASUALTIES")

casualtiesVsType = casualtiesVsType[which(casualtiesVsType$CASUALTIES > 15),]

```

```{r plotCasualtiesVsType, fig.height=6}

barchart(CASUALTIES ~ EVTYPE,
         data = casualtiesVsType,
         scales = list(x=list(rot=45)),
         ylim = range(casualtiesVsType$CASUALTIES, na.rm = TRUE)
         )

```

```{r calculateDollars}

calculateDollars <- function(base, exponent) {
  dollars = NA
  
  if (exponent == "K") {
    dollars = base * 1000
  } else if (exponent == "M") {
    dollars = base * 1000000
  } else if (exponent == "B") {
    dollars = base * 1000000000
  } else if (exponent == "H") {
    dollars = base * 100
  } else {
    dollars = base * 1000
  }
  dollars
}

```

```{r calculateDollarDamages}

simpleData <- cbind(simpleData, PROPDOLLARS = NA, CROPDOLLARS = NA, TOTALDOLLARS = NA)

simpleData$PROPDOLLARS <- calculateDollars(simpleData$PROPDMG, simpleData$PROPDMGEXP)

simpleData$CROPDOLLARS <- calculateDollars(simpleData$CROPDMG, simpleData$CROPDMGEXP)

simpleData$TOTALDOLLARS <- simpleData$PROPDOLLARS + simpleData$CROPDOLLARS
simpleData$TOTALDOLLARSK <- simpleData$TOTALDOLLARS / 1000

```

```{r totalDamages}

damageVsType <- aggregate(simpleData$TOTALDOLLARSK, by=list(simpleData$EVTYPE), FUN=mean, na.rm=TRUE)

names(damageVsType) = c("EVTYPE", "TOTALDOLLARSK")

damageVsType = damageVsType[which(damageVsType$TOTALDOLLARSK > 500),]

```

```{r plotDamagesVsType, fig.height=6}

barchart(TOTALDOLLARSK ~ EVTYPE,
         data = damageVsType,
         scales = list(x=list(rot=45)),
         ylim = range(damageVsType$TOTALDOLLARSK, na.rm = TRUE)
         )

```

## Results
### Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?

Heat waves have the highest mean number of casualties (where casualties are defined as either injuries or deaths). 

### Across the United States, which types of events have the greatest economic consequences?




